Computational Methods for Principle Component Analysis Applied to Genotype Data

### 2018-02-26 doc/paper.tex
Change test_hdpcaRand time to runtimes_sum[projRand] (rather than runtimes[projRand])
Create comparison tables and rsvd refs in paper.tex

\subsubsection{Randomized SVD}
In addition to estimating the PC scores of the study individuals,
we have another algorithm that aims at improving the decomposition of the
training data.
For SP, AP, ADT, and OADT, we need to find the covariance matrix of the training
data, which has dimension $n \times n$, and conduct eigendecomosition on it.
However, when $n$ is exceedingly large,
the covariance matrix cannot be loaded into the memory of an average personal
computer.
A randomized SVD (RSVD)algorithm has been developed to handle this issue.
RSVD is slower than standard SVD,
but it does not require loading the whole covariance matrix into the memory.

@@

@@

@@

\begin{table} 
  \centering
  \begin{tabular}{|l|l|}
    \hline
    Method & Computational Complexity \\ 
    \hline
    Simple Projection (SP) & $\bO[n^2p + mk_1p]$ \\
    \hline
    Augment-Decompose-Transform (ADT) & $\bO[n^2 p + mn(p + n^2)]$ \\
    \hline
    Adjusted Projection (AP) & $\bO[n^2p + mk_1p]$ \\
    \hline
    Online Augment-Decompose-Transform (OADT) & $\bO[n^2 p + mk_3(p + k_3 n)]$ \\
    \hline
  \end{tabular}
  \caption{
    Comparison of computational complexity for the three methods.
    Here $p$ is the number of SNPs,
    $n$ is the number of individuals in the reference group,
    and $m$ is the number of individuals in the  study group.
    For projection, $k_1$ is the number of PCs needed
    For OADT, $k_3$ is the number of PCs calculated by online SVD, which is usually much less than $n$.}
  \label{tbl:cplx}
\end{table}

\begin{table} 
  \centering
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    Method & Reference Runtimes & Study Runtimes & Reference Center Error & Golden Standard Error \\ 
    \hline
    Simple Projection (SP) &  \\
    \hline
    Augment-Decompose-Transform (ADT) & $\bO[n^2 p + mn(p + n^2)]$ \\
    \hline
    Adjusted Projection (AP) & $\bO[n^2p + mk_1p]$ \\
    \hline
    Online Augment-Decompose-Transform (OADT) & $\bO[n^2 p + mk_3(p + k_3 n)]$ \\
    \hline
  \end{tabular}
  \caption{
    Comparison of computational complexity for the three methods.
    Here $p$ is the number of SNPs,
    $n$ is the number of individuals in the reference group,
    and $m$ is the number of individuals in the  study group.
    For projection, $k_1$ is the number of PCs needed
    For OADT, $k_3$ is the number of PCs calculated by online SVD, which is usually much less than $n$.}
  \label{tbl:cplx}
\end{table}

>>>>

\subsubsection{Randomized SVD}
In addition to estimating the PC scores of the study individuals,
we have another algorithm that aims at improving the decomposition of the
training data.
For SP, AP, ADT, and OADT, we need to find the covariance matrix of the training
data, which has dimension $n \times n$, and conduct eigendecomosition on it.
However, when $n$ is exceedingly large,
the covariance matrix cannot be loaded into the memory of an average personal
computer.
A randomized SVD (RSVD) algorithm has been developed to handle this issue.
RSVD is slower than standard SVD,
but it does not require loading the whole covariance matrix into the memory.
For the details of the RSVD algorithm, see \cite{halko} and \cite{abraham}.

@@

\bibitem[Abraham \& Inouye(2014)]{abraham} Abraham, G., \& Inouye, M. (2014). Fast principal component analysis of large-scale genome-wide data. {\it PloS one}, 9(4), e93766.

@@

\bibitem[Halko et al.(2011)]{halko} Halko, N., Martinsson, P. G., \& Tropp, J. A. (2011). Finding structure with randomness: Probabilistic algorithms for constructing approximate matrix decompositions. {\it SIAM review}, 53(2), 217-288.

@@

\begin{table} 
  \centering
  \begin{tabular}{|l|l|l|}
    \hline
    Method & Reference Complexity & Study Complexity \\ 
    \hline
    Simple Projection (SP) & $\bO[n^2p]$ & $\bO[mkp]$ \\
    \hline
    Randomized Simple Projection (RSP) & $\bO[k^2p]$ & $\bO[mkp]$ \\
    \hline
    Augment-Decompose-Transform (ADT) & $\bO[n^2 p]$ & $\bO[mn(p + n^2)]$ \\
    \hline
    Adjusted Projection (AP) & $\bO[n^2p]$ &  $\bO[mkp]$ \\
    \hline
    Randomized Adjusted Projection (RAP) & $\bO[k^2p]$ &  $\bO[mkp]$ \\
    \hline
    Online Augment-Decompose-Transform (OADT) & $\bO[n^2 p]$ & $\bO[mk(p + k n)]$ \\
    \hline
    Randomized Online Augment-Decompose-Transform (ROADT) & $\bO[k^2 p]$ & $\bO[mk(p + k n)]$ \\
    \hline
  \end{tabular}
  \caption{
    Comparison of computational complexity for different methods.
    Here $p$ is the number of SNPs,
    $n$ is the number of individuals in the reference group,
    $m$ is the number of individuals in the  study group,
    and $k$ is the number of PCs calculated by the corresponding method.
  }
  \label{tbl:cplx}
\end{table}

\begin{table} 
  \centering
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    Method & Reference Runtimes & Study Runtimes & Golden Standard Error \\ 
    \hline
    SP & 411 & 425 & 1.95 \\
    \hline
    RSP & 2504 & 241 & 1.95 \\
    \hline
    ADT & 411 & 29813 & 0.00 \\
    \hline
    AP & 411 & 482 1 & 1.12 \\
    \hline
    RAP & 2505 & 250 & 1.83 \\ 
    \hline
    OADT & 2518 & 103 & 0.40 \\
    \hline
    ROADT & 2518 & 103 & 1.93 \\
    \hline
  \end{tabular}
  \caption{
    Comparison of runtimes and accuracy for different methods applied to the real-life datasets.
    The reference individuals are from the 1000 Genomes dataset with 2492 unrelated individuals.
    The study sample contain 1000 individuals from the UKBioBank dataset.
    There are about 125,000 loci shared by the reference and study data.
    The golden standard error is the square root of the averate squared distance between the study individuals' PC scores predicted by different methods compared to the golden standard result, which in our case is the PC scores predicted by ADT.
  }
  \label{tbl:cplx}
\end{table}
